---
title: "Lab2"
author: "Harshini Karthikeyan"
date: 'April 10th 2025'
format: html
output:
  html_document:
    number_sections: yes

    code-fold: true
    code-summary: "Show code"
    backgroundcolor: whitesmoke
    monobackgroundcolor: lightsteelblue
    fontcolor: steelblue
    embed-resources: true
  pdf_document: default
mainfont: "Times New Roman"
editor: visual
toc: true
toc_float: true
embed-resources: true
---

# Lab 2

```{r}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(readxl)
library(ggrepel)
library(scales)
library(leaflet)
```

## Part 1:

1.  **While there are certainly issues with this image, do your best to tell the story of this graph in words. That is, what is this graph telling you? What do you think the authors meant to convey with it?**

I believe this plot is trying to show us the proportion of people in a particular country that believe a vaccine is safe. It shows us one observation per country where the y-axis appears to be country? It also sorted the observations by proportion, so that it appears that the proportions increase as the y-axis goes up.

2.  **List the variables that appear to be displayed in this visualization. *Hint: Variables refer to columns in the data.***

The country, the global region of the country and the percentage of people in the country that believe a vaccine is safe.

3.  **Now that you’re versed in the grammar of graphics (e.g., `ggplot`), list the *aesthetics* used and which *variables* are mapped to each.**

The x-axis the percentage of people in the country that believe a vaccine is safe. The y-axis is the country, the color is filled by global region.

4.  **What type of graph would you call this? Meaning, what `geom` would you use to produce this plot?**

I would call this a scatter plot and would use geom point to produce it.

5.  **Provide at least four problems or changes that would improve this graph. *Please format your changes as bullet points!***

-   I think this is a graph that is faceted by region of the world, but it lends to a strange stacked graph like view that is not very intuitive, and leaves some plots stranded from their axes. I would try faceting differently or trying to remove the facet entirely.

-   I think having the y-axis be the countries is a bit strange, and that sorting it by the proportion feels odd, as though it is increasing over time or something, I would try sorting it alphabetically or removing the y-axis and making it a box-plot instead.

-   I think having the legend at the bottom is unnecessary as the global region is also written directly onto each facet of the plot, so I would remove one of those, preferably the legend, depending on how my plot turns out.

-   I think having to state that the dark vertical lines represent region medians but also not writing the number of each on the plot itself is confusing and forces the reader to then look for th median by finding the axis

## **Improving the Bad Visualization**

6.  **Improve the visualization above by either re-creating it with the issues you identified fixed OR by creating a new visualization that you believe tells the same story better.**

```{r}
data<-read_excel("C:\\Users\\Geetha\\Downloads\\wgm2018-dataset-crosstabs-all-countries.xlsx")
head(data)
```

```{r}
View(data)
```

```{r}
national <- data[c('...1', '...2', '...3', 'National results','...5')]
colnames(national) <- c("Country", "Question", "Response", "National_Results_Col_Per", "National_Results_Count")

national <- national[!(national$Country %in% c(NA, 'Country')),]
#drop title row since renamed cols alr

```

```{r}
national$QNumber <- str_split_i(national$Question, " ", 1)
national
```

```{r}

asia <- c('Afghanistan', 'Bangladesh', 'Cambodia', 'China', 'India', 'Indonesia', 'Japan', 'Laos', 'Malaysia', 'Mongolia', 'Myanmar', 'Nepal', 'Pakistan', 'Philippines', 'Singapore', 'South Korea', 'Sri Lanka', 'Taiwan', 'Thailand', 'Vietnam'
)

mena <- c('Algeria', 'Egypt', 'Iran', 'Iraq', 'Israel', 'Jordan', 'Kuwait', 'Lebanon','Libya', 'Morocco', 'Palestine', 'Saudi Arabia', 'Tunisia', 'United Arab Emirates', 'Yemen'
)

americas <- c('Argentina', 'Bolivia', 'Brazil', 'Canada', 'Chile', 'Colombia', 'Costa Rica', 'Dominican Republic', 'Ecuador', 'El Salvador', 'Guatemala', 'Haiti', 'Honduras','Mexico', 'Nicaragua', 'Panama', 'Paraguay', 'Peru', 'United States', 'Uruguay', 'Venezuela'
)

sub_sahara <- c('Benin', 'Botswana', 'Burkina Faso', 'Burundi', 'Cameroon', 'Congo, Rep.','Comoros', 'Chad','Eswatini','Ethiopia', 'Gabon', 'Ghana', 'Guinea', 'Ivory Coast', 'Kenya', 'Liberia', 'Madagascar', 'Malawi', 'Mali', 'Mauritania', 'Mauritius','Mozambique', 'Namibia', 'Niger', 'Nigeria', 'Rwanda', 'Senegal','Sierra Leone', 'South Africa', 'Tanzania', 'The Gambia', 'Togo','Uganda', 'Zambia', 'Zimbabwe'
)

europe <- c('Albania', 'Austria', 'Belgium', 'Bosnia and Herzegovina', 'Bulgaria','Croatia', 'Cyprus', 'Czech Republic', 'Denmark', 'Finland', 'France','Germany', 'Greece', 'Hungary', 'Iceland', 'Ireland', 'Italy', 'Kosovo','Luxembourg', 'Macedonia', 'Malta', 'Netherlands', 'Norway', 'Poland', 'Northern Cyprus', 'Montenegro', 'Portugal', 'Romania', 'Serbia', 'Slovakia', 'Slovenia', 'Spain', 'Sweden', 'Switzerland', 'Turkey', 'United Kingdom'
)

former_soviet <- c('Armenia', 'Azerbaijan', 'Belarus', 'Estonia', 'Georgia', 'Kazakhstan', 'Kyrgyzstan', 'Latvia', 'Lithuania', 'Moldova', 'Russia', 'Tajikistan', 'Ukraine', 'Uzbekistan', 'Turkmenistan'
)

oceania <- c('Australia', 'New Zealand')
```

```{r}
national1 <- national |>
  fill(QNumber, .direction = "down") |>
  #Source for .direction: https://tidyr.tidyverse.org/reference/fill.html
  filter(QNumber == "Q25") |>
  filter(Response == "Strongly agree" | Response == "Somewhat agree") |> group_by(Country) |>
  summarize(Agree_Percent = sum(as.numeric(National_Results_Col_Per)), .groups = "drop") |> mutate(Region = case_when(
    
    Country %in% asia ~ "Asia",
    Country %in% mena ~ "Middle East and North Africa",
    Country %in% americas ~ "Americas",
     Country %in% sub_sahara ~ "Sub-Saharan Africa",
    Country %in% europe ~ "Europe",
    
    Country %in% former_soviet ~ "Former Soviet Union",
    
    Country %in% oceania ~ "Oceania",
    
    TRUE ~ "Other" #used to check, there are none that fall in this category
  ))
  #summarize function : https://stackoverflow.com/questions/62891736/sum-sub-groups-with-dplyr




```

```{r}

outliers <- national1|>
  group_by(Region)|>
  mutate(
    Q1 = quantile(Agree_Percent, 0.25),  Q3 = quantile(Agree_Percent, 0.75),
    IQR = Q3 - Q1,
    is_outlier = Agree_Percent < (Q1 - 1.5 * IQR) | Agree_Percent > (Q3 + 1.5 * IQR)
  ) |>
  filter(is_outlier)
```

```{r}

ggplot(national1, aes(x = Region,y=Agree_Percent,fill = Region))+
  geom_boxplot(outlier.shape= NA,alpha = 0.7) +  
  geom_point(data = outliers, aes(x =Region, y= Agree_Percent), shape =16, size =3, color = "black") +
  geom_text(
    data =outliers,
    aes(label= Country), vjust =0.25, hjust=-0.25,  size = 4, family="serif"
  )+ scale_y_continuous(labels = percent_format(accuracy = 1)) +
  labs(
    title = "The Percentage of People Who Believe Vaccines are Safe by Global Region",
    x= "Region",y= NULL,
  ) +theme_minimal(base_family = "serif")+
  
  theme(axis.text.x = element_text(angle = 20, hjust = 0.75), axis.text = element_text(size = 10,color = "black"), legend.position = "none"
        )


```

# **Part Two: Broad Visualization Improvement**

## **Second Data Visualization Improvement**

7.  **Select a data visualization in the report that you think could be improved. Be sure to cite both the page number and figure title. Do your best to tell the story of this graph in words. That is, what is this graph telling you? What do you think the authors meant to convey with it?**

Chart 5.7: Map of proportions of people reporting vaccinating their children on page 121. This graph is telling us what percentage of the people in a particular country say their children have been vaccinated. It compares between countries, highlightng where this percentage is particularly high or low.

8.  **List the variables that appear to be displayed in this visualization.**

Percentage of people who answered ‘yes’ to their children having received vaccines and Country,

9.  **Now that you’re versed in the grammar of graphics (ggplot), list the aesthetics used and which variables are specified for each.**

The map shows the variable country, the color is filled by the percentage of people who answered 'yes'.

10. **What type of graph would you call this?**

I would call this a chloropleth map!

#Source: <https://support.esri.com/en-us/gis-dictionary/choropleth-map#:~:text=From%20the%20Greek%20terms%20choro,combining%20different%20sets%20of%20symbols.>

11. **List all of the problems or things you would improve about this graph.**

-   I would make the color scaling more distinct and more apparent as it is difficult to visibly see the percentage or people who said their children had been vaccinated.

-   I might add in global region to see if that shows any distinctions.

-   I would make the countries that were not surveyed a more ditinct color from the background color as it is hard to notice island nations and the like.

12. **Improve the visualization above by either re-creating it with the issues you identified fixed OR by creating a new visualization that you believe tells the same story better.**

```{r}
national2 <- national |>
  fill(QNumber, .direction = "down") |>
  #Source for .direction: https://tidyr.tidyverse.org/reference/fill.html
  filter(QNumber == "Q28")|> filter(Response == "Yes") |> mutate(Region = case_when(
    
    Country %in% asia ~ "Asia",
    Country %in% mena ~ "Middle East and North Africa",
    Country %in% americas ~ "Americas",
     Country %in% sub_sahara ~ "Sub-Saharan Africa",
    Country %in% europe ~ "Europe",
    
    Country %in% former_soviet ~ "Former Soviet Union",
    
    Country %in% oceania ~ "Oceania",
    
    TRUE ~ "Other" #used to check, there are none that fall in this category
  ))
head(national2)
```

```{r}
library(rnaturalearth)
library(sf)
library(rnaturalearthdata)

world <- ne_countries(scale = "medium", returnclass = "sf")
world$Country <- tolower(world$name_long)
national2$Country <- tolower(national2$Country)

world_data <- left_join(world, national2, by = c("Country" = "Country"))
world_data$National_Results_Col_Per <- as.numeric(world_data$National_Results_Col_Per)
```

```{r}
pal <- colorBin(
  palette ="YlGnBu",
  domain = world_data$National_Results_Col_Per,
  bins = c(0, 0.70, 0.80, 0.90, 0.95, 1.00), na.color ="#d3d3d3"  
)

leaflet(world_data) |> addTiles() |>
  addPolygons(
    fillColor = ~pal(National_Results_Col_Per), weight = 1,
    color = "white", fillOpacity = 0.8,
    label = ~paste0(Country, ": ", National_Results_Col_Per, "%"),
    highlight = highlightOptions(
      weight = 2, color = "#666",
      fillOpacity = 0.9, bringToFront = TRUE
    )
    #Source: https://r-charts.com/spatial/interactive-maps-leaflet/
  ) |>
  
  addLegend(
    pal = pal, values = ~National_Results_Col_Per,
    title = "Proportions of people", position = "bottomright"
  ) 

```

## **Third Data Visualization Improvement**

```{r}

national3 <- data[c('...1', '...2', '...3',
                    'National results','...5',
                    'Gender','...7', '...8', '...9',
                    'Age Cohort', '...11', '...12',
                    '...13', '...14', '...15',
                    'General Educational Background', 
                    '...17', '...18', '...19', 
                    '...20', '...21', 'Area Type', 
                    '...23', '...24', '...25')]

colnames(national) <- c("Country", "Question", "Response",
                        "National_Results_Col_Per", "National_Results_Count",
                        "Man_Per", "Man_num", "Woman_Per", "Woman_num",
                        "15-29_Per", "15-29_Num", "15-29_Per",
                        "15-29_Per", "15-29_Per", "15-29_Per")

national <- national[!(national$Country %in% c(NA, 'Country')),]
national$QNumber <- str_split_i(national$Question, " ", 1)

national1 <- national |> filter(Country == 'France') |>
  fill(QNumber, .direction = "down") |>
  #Source for .direction: https://tidyr.tidyverse.org/reference/fill.html
  filter(QNumber == "Q25") |>
  filter(Response == "Strongly disagree" | Response == "Somewhat disagree") |> group_by(Country) |>
  summarize(Agree_Percent = sum(as.numeric(National_Results_Col_Per, )), .groups = "drop") |> mutate(Region = case_when(
    
    Country %in% asia ~ "Asia",
    Country %in% mena ~ "Middle East and North Africa",
    Country %in% americas ~ "Americas",
     Country %in% sub_sahara ~ "Sub-Saharan Africa",
    Country %in% europe ~ "Europe",
    
    Country %in% former_soviet ~ "Former Soviet Union",
    
    Country %in% oceania ~ "Oceania",
    
    TRUE ~ "Other" #used to check, there are none that fall in this category
  ))
  #summarize function : https://stackoverflow.com/questions/62891736/sum-sub-groups-with-dplyr



```
