---
title: "Lab 3"
date: 'April 21st 2025'
format: 
  dashboard:
    theme: united
    orientation: columns
    nav-buttons: 
      - icon: github
        href: https://github.com/hersheyk/STAT541/tree/main
        aria-label: GitHub
      - icon: linkedin
        href: https://www.linkedin.com/in/harshini-karthikeyan-9512ab1b0/
        aria-label: LinkedIn
      - icon: instagram
        href: https://www.instagram.com/harshini.825/
        aria-label: Instagram
logo: images/Wellcome_Trust_logo.svg.png
logo-alt: Logo of Wellcome black with letter 'w' in the middle
editor: visual
embed-resources: true
---

```{r libraries}
#| message: false
library(tidyverse)
library(dplyr)
library(ggplot2)
library(readxl)
library(ggrepel)
library(scales)
library(leaflet)
library(knitr)
library(rnaturalearth)
library(sf)
library(rnaturalearthdata)

```

```{r read in data}
#| message: false
data<-read_excel("C:\\Users\\Geetha\\Downloads\\wgm2018-dataset-crosstabs-all-countries.xlsx")
#head(data)
```

```{r subset data}
#| message: false
national <- data[c('...1', '...2', '...3', 'National results','...5')]
colnames(national) <- c("Country", "Question", "Response", "National_Results_Col_Per", "National_Results_Count")

national <- national[!(national$Country %in% c(NA, 'Country')),]
#drop title row since renamed cols alr

```

```{r split to get question number}
#| message: false
national$QNumber <- str_split_i(national$Question, " ", 1)
#national
```

```{r name global region vectors}

asia <- c('Afghanistan', 'Bangladesh', 'Cambodia', 'China', 'India', 'Indonesia', 'Japan', 'Laos', 'Malaysia', 'Mongolia', 'Myanmar', 'Nepal', 'Pakistan', 'Philippines', 'Singapore', 'South Korea', 'Sri Lanka', 'Taiwan', 'Thailand', 'Vietnam'
)

mena <- c('Algeria', 'Egypt', 'Iran', 'Iraq', 'Israel', 'Jordan', 'Kuwait', 'Lebanon','Libya', 'Morocco', 'Palestine', 'Saudi Arabia', 'Tunisia', 'United Arab Emirates', 'Yemen'
)

americas <- c('Argentina', 'Bolivia', 'Brazil', 'Canada', 'Chile', 'Colombia', 'Costa Rica', 'Dominican Republic', 'Ecuador', 'El Salvador', 'Guatemala', "Haiti", 'Honduras','Mexico', 'Nicaragua', 'Panama', 'Paraguay', 'Peru', 'United States', 'Uruguay', 'Venezuela'
)

sub_sahara <- c('Benin', 'Botswana', 'Burkina Faso', 'Burundi', 'Cameroon', 'Congo, Rep.','Comoros', 'Chad','Eswatini','Ethiopia', 'Gabon', 'Ghana', 'Guinea', 'Ivory Coast', 'Kenya', 'Liberia', 'Madagascar', 'Malawi', 'Mali', 'Mauritania', 'Mauritius','Mozambique', 'Namibia', 'Niger', 'Nigeria', 'Rwanda', 'Senegal','Sierra Leone', 'South Africa', 'Tanzania', 'The Gambia', 'Togo','Uganda', 'Zambia', 'Zimbabwe'
)

europe <- c('Albania', 'Austria', 'Belgium', 'Bosnia and Herzegovina', 'Bulgaria','Croatia', 'Cyprus', 'Czech Republic', 'Denmark', 'Finland', 'France','Germany', 'Greece', 'Hungary', 'Iceland', 'Ireland', 'Italy', 'Kosovo','Luxembourg', 'Macedonia', 'Malta', 'Netherlands', 'Norway', "Poland", 'Northern Cyprus', 'Montenegro', 'Portugal', 'Romania', 'Serbia', 'Slovakia', 'Slovenia', 'Spain', 'Sweden', 'Switzerland', 'Turkey', 'United Kingdom'
)

former_soviet <- c('Armenia', 'Azerbaijan', 'Belarus', 'Estonia', 'Georgia', 'Kazakhstan', 'Kyrgyzstan', 'Latvia', 'Lithuania', 'Moldova', 'Russia', 'Tajikistan', 'Ukraine', 'Uzbekistan', 'Turkmenistan'
)

oceania <- c('Australia', 'New Zealand')
```

# Children Vaccination Statistics

## Plots {.tabset}

```{r create nat2}

#| message: false
national2 <- national |>
  fill(QNumber, .direction = "down") |>
  #Source for .direction: https://tidyr.tidyverse.org/reference/fill.html
  filter(QNumber == "Q28")|> filter(Response == "Yes") |> mutate(Region = case_when(
    
    Country %in% asia ~ "Asia",
    Country %in% mena ~ "Middle East and North Africa",
    Country %in% americas ~ "Americas",
     Country %in% sub_sahara ~ "Sub-Saharan Africa",
    Country %in% europe ~ "Europe",
    Country %in% former_soviet ~ "Former Soviet Union",
    
    Country %in% oceania ~ "Oceania",
    
    TRUE ~ "Other" #used to check, there are none that fall in this category
  ))
#head(national2)

```

```{r connect nat2 to world dataset}

world <- ne_countries(scale = "medium", returnclass = "sf")
world$Country <- tolower(world$name_long)
national2$Country <- tolower(national2$Country)

world_data <- left_join(world, national2, by = c("Country" = "Country"))
world_data$National_Results_Col_Per <- as.numeric(world_data$National_Results_Col_Per)


#colorpalette
pal <- colorBin(
  palette =c("#ffffd4", "#fec44f", "#fe9929", "#fc4e2a", "#b10026"),
  #https://loading.io/color/feature/YlOrRd-8
  domain = world_data$National_Results_Col_Per,
  bins =c(0, 0.70, 0.80, 0.90, 0.95, 1.00), na.color ="#d3d3d3"  
)

```

### The percentage of vaccinated children by country

```{r plot nat2}
#| title: The percentage of vaccinated children by region
leaflet(world_data) |> addTiles() |>
  addPolygons(
    fillColor = ~pal(National_Results_Col_Per), weight = 1,
    color = "white", fillOpacity = 0.8,
    label = ~paste0(Country, ": ", National_Results_Col_Per, "%"),
    highlight = highlightOptions(
      weight = 2, color = "#666",
      fillOpacity = 0.9, bringToFront = TRUE
    )
    #Source: https://r-charts.com/spatial/interactive-maps-leaflet/
  ) |>
  
  addLegend(
    pal = pal, values = ~National_Results_Col_Per,
    title = "Percentage of people who say  <br> their child has been vaccinated", position = "bottomright"
    #Source line break: https://www.reddit.com/r/Rlanguage/comments/6bsji1/add_line_break_to_leaflet_pop_up/
  ) 

```

```{r numeric}
national2$National_Results_Col_Per <- as.numeric(national2$National_Results_Col_Per)
```

### The percentage of vaccinated children by region

```{r plot box}


ggplot(national2, aes(x = Region,y=National_Results_Col_Per,fill = Region))+
  geom_boxplot(outlier.shape= NA,alpha = 0.7) +  
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  scale_fill_manual(values = c(
    "Sub-Saharan Africa" = "#581209",
    "Europe" = "#e31a1c",
    "Asia" = "#CC5500",
    "Americas" = "#e88d13",
    "Middle East and North Africa" = "#e1a612",
    "Oceania" = "#fb760b",
    "Former Soviet Union" = "#c34328")) +
  labs(
    title = "The Percentage of Vaccinated Children by Global Region",
    x= "Region",y= NULL,
  ) +theme_minimal(base_family = "serif")+
  
  theme(axis.text.x = element_text(angle = 20, hjust = 0.75), axis.text = element_text(size = 10,color = "black"), legend.position = "none"
        )

```

## More Numbers

```{r}
#| label: calculate-values
lowest_country <- national2 |>
  filter(National_Results_Col_Per == min(National_Results_Col_Per, na.rm = TRUE)) |>
  distinct(Country) |>
  pull(Country)

highest_country <- national2 |>
  filter(National_Results_Col_Per == max(National_Results_Col_Per, na.rm = TRUE)) |>
  distinct(Country) |>
  pull(Country)


```

### Lowest

```{r values}
#| content: valuebox
#| title: "Country with Lowest Percentage of Vaccinated Children"
#| color: danger
list(
  value = toupper(lowest_country)
)
```


### Table

```{r table}

national2 |>
  select(Country, Region, National_Results_Col_Per) |> mutate(National_Results_Col_Per = scales::percent(National_Results_Col_Per, accuracy = 1)) |>
  arrange(desc(National_Results_Col_Per)) |> head(10) |>
  kable(col.names = c("Country", "Region", "% Believe Vaccines Are Safe"))
```

# Belief in Vaccine Safety Statistics

```{r nat3}

national3 <- national |>
  fill(QNumber, .direction = "down") |>
  #Source for .direction: https://tidyr.tidyverse.org/reference/fill.html
  filter(QNumber == "Q25")|> filter(Response %in% c("Strongly agree", "Somewhat agree")) |> group_by(Country) |> summarize(Agree_Percent = sum(as.numeric(National_Results_Col_Per)), .groups = "drop") |> mutate(Region = case_when(
    
    Country %in% asia ~ "Asia",
    Country %in% mena ~ "Middle East and North Africa",
    Country %in% americas ~ "Americas",
     Country %in% sub_sahara ~ "Sub-Saharan Africa",
    Country %in% europe ~ "Europe",
    Country %in% former_soviet ~ "Former Soviet Union",
    
    Country %in% oceania ~ "Oceania",
    
    TRUE ~ "Other" #used to check, there are none that fall in this category
  ))
```

```{r nat3 to world}
world <- ne_countries(scale = "medium", returnclass = "sf")
world$Country <- tolower(world$name_long)
national3$Country <- tolower(national3$Country)

world_data2 <- left_join(world, national3, by = c("Country" = "Country"))
world_data2$National_Results_Col_Per <- as.numeric(world_data$National_Results_Col_Per)

```

## Safety Plots {.tabset}

### The Percentage of People who Believe Vaccines are Safe by Country

```{r plot nat3}
#| title: The percentage of people who believe vaccines are safe by Country
leaflet(world_data2) |> addTiles() |>
  addPolygons(
    fillColor = ~pal(Agree_Percent), weight = 1,
    color = "white", fillOpacity = 0.8,
    label = ~paste0(Country, ": ", Agree_Percent, "%"),
    highlight = highlightOptions(
      weight = 2, color = "#666",
      fillOpacity = 0.9, bringToFront = TRUE
    )
    #Source: https://r-charts.com/spatial/interactive-maps-leaflet/
  ) |>
  
  addLegend(
    pal = pal, values = ~ Agree_Percent,
    title = "Percentage of people who believe <br> vaccines are safe", position = "bottomright"
    #Source line break: https://www.reddit.com/r/Rlanguage/comments/6bsji1/add_line_break_to_leaflet_pop_up/
  ) 

```
### The Percentage of People who Believe Vaccines are Safe by Country
```{r}

ggplot(national3, aes(x = Region,y= Agree_Percent,fill = Region))+
  geom_boxplot(outlier.shape= NA,alpha = 0.7) +  
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  scale_fill_manual(values = c(
    "Sub-Saharan Africa" = "#581209",
    "Europe" = "#e31a1c",
    "Asia" = "#CC5500",
    "Americas" = "#e88d13",
    "Middle East and North Africa" = "#e1a612",
    "Oceania" = "#fb760b",
    "Former Soviet Union" = "#c34328")) +
  labs(
    title = "The Percentage of People Who Believe Vaccines are Safe by Global Region",
    x= "Region",y= NULL,
  ) +theme_minimal(base_family = "serif")+
  
  theme(axis.text.x = element_text(angle = 20, hjust = 0.75), axis.text = element_text(size = 10,color = "black"), legend.position = "none"
        )

```
```{r}
#| label: calculate-vals for national3
lowest_country <- national3 |>
  filter(Agree_Percent == min(Agree_Percent, na.rm = TRUE)) |>
  distinct(Country) |>
  pull(Country)

highest_country <- national3 |>
  filter(Agree_Percent == max(Agree_Percent, na.rm = TRUE)) |>
  distinct(Country) |>
  pull(Country)


```

## Numbers
### Lowest {height = "60%}

```{r values2}
#| content: valuebox
#| title: "Country with Lowest Belief in Vaccine Safety"
#| color: danger
list(
  value = toupper(lowest_country)
)
```

### Highest {height = "60%}
```{r}
#| content: valuebox
#| title: "Country with Highest Belief in Vaccine Safety"
#| color: danger
list(
  value = toupper(highest_country)
)
```

### Nat 3
```{r}

national3 |>
  select(Country, Region, Agree_Percent) |> mutate(Agree_Percent = scales::percent(Agree_Percent, accuracy = 1)) |>
  arrange(desc(Agree_Percent)) |> head(10) |>
  kable(col.names = c("Country", "Region", "% Believe Vaccines Are Safe"))
```

